---
title: "data-analisys"
author: "Mitja Bof"
date: "10/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Loading the Manifesto project data
Data is available [here](https://manifesto-project.wzb.eu/datasets), codebook available [here](https://manifesto-project.wzb.eu/down/data/2022a/codebooks/codebook_MPDataset_MPDS2022a.pdf)

Data cleaning and preparation

```{r}
mp <- read_csv("data/MPDataset_MPDS2022a.csv")

# Filtering countries of interest
# Selecting variables of interest
countries <- c("Sweden", "Norway", "Denmark", "Finland", "Netherlands",
               "France", "Italy", "Spain", "Germany", "United Kingdom",
               "Ireland")
vars <- c("countryname", "edate", "partyname", "parfam", "pervote",
          "absseat", "totseats", "peruncod")
mp <- mp |> filter(countryname %in% countries) |> 
  select(vars, per101:per706) |> 
  mutate(perseat = absseat / totseats) |> filter(perseat > 0.02)

```

For every country, for every election, calculate within group variability.
WGV: Sum of squared distance from the mean of each point in the n-dimensional space.

```{r}
elections <- mp |> group_by(countryname, edate) |> summarize(parties = n())
el_iterable <- seq_along(elections$parties)

for (e in el_iterable) {
  sub_df <- mp |> 
    filter(countryname == elections$countryname[e], 
           edate == elections$edate[e])
  center <- sub_df |> 
    summarize(across(per101:per706, mean))
  wgv_squared <- 0
  for (p in 1:elections$parties[e]) {
    party <- sub_df[p, 9:64]
    distance <- (party - center)^2 |> sum()
    wgv_squared <- wgv_squared + distance
  }
  elections$wgv[e] <- sqrt(wgv_squared / elections$parties[e])
}
```

Exploratory analysis of WGV (partisan heterogeneity):
There seems to be a positive relationship between number of party in an election and my measure of party differences. Is this the result of poor operationalization or is it a substantive finding?

```{r}
elections |> ggplot(aes(x = parties, y = wgv)) + 
  geom_point() + geom_smooth(method = "lm")

elections |> group_by(parties) |> summarize(mean = mean(wgv), n = n()) |>
  ggplot(aes(x = parties, y = mean)) + geom_col() + 
  geom_text(aes(label = n), vjust = -0.2)
```

Effective number of parties
No of partie further than x from mean.
Weight dist by votes
Get map visualization

```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

european_union <- c("Austria","Belgium","Bulgaria","Croatia","Cyprus",
                    "Czech Rep.","Denmark","Estonia","Finland","France",
                    "Germany","Greece","Hungary","Ireland","Italy","Latvia",
                    "Lithuania","Luxembourg","Malta","Netherlands","Poland",
                    "Portugal","Romania","Slovakia","Slovenia","Spain",
                    "Sweden","United Kingdom")

map <- ne_countries(scale = 50, returnclass = 'sf') |> 
  filter(name %in% european_union)
bbox_europe <- st_bbox(c(xmin = -10, ymin = 20, xmax = 50, ymax = 80), 
                       crs = st_crs(map))
elections_mean <- elections |> group_by(countryname) |>
  summarize(wgv = mean(wgv))
map_cropped <- st_crop(map, bbox_europe) |> 
  left_join(elections_mean, by = c("name" = "countryname"))

ggplot(data = map_cropped) +
  geom_sf(mapping = aes(fill = wgv)) +
  scale_fill_gradient(name = "Party differences", 
                      low = "#FF0000FF", high = "#FFFF00FF", 
                      na.value = "grey50") +
  labs(title = "Party heterogeneity in Europe") +
  theme(plot.title.position = "plot")

```

Loading electoral system data

```{r}
syst <- read_csv("data/elect_systems_national_leg_europe.csv") |> 
  filter(Country %in% countries)
```

