---
title: "data-analisys"
author: "Mitja Bof"
date: "10/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(infer)
```

Loading the Manifesto project data
Data is available [here](https://manifesto-project.wzb.eu/datasets), codebook available [here](https://manifesto-project.wzb.eu/down/data/2022a/codebooks/codebook_MPDataset_MPDS2022a.pdf)

Data cleaning and preparation

```{r}
mp <- read_csv("data/MPDataset_MPDS2022a.csv")

# Filtering countries of interest
# Selecting variables of interest
countries <- c("Sweden", "Norway", "Denmark", "Finland", "Netherlands",
               "France", "Italy", "Spain", "Germany", "United Kingdom",
               "Ireland")
vars <- c("countryname", "edate", "partyname", "parfam", "pervote",
          "absseat", "totseats", "peruncod")
mp <- mp |> filter(countryname %in% countries) |> 
  select(vars, per101:per706) |> 
  mutate(perseat = absseat / totseats) |> filter(perseat > 0.02)

```

For every country, for every election, calculate within group variability.
WGV: Sum of squared distance from the mean of each point in the n-dimensional space.

```{r}
elections <- mp |> group_by(countryname, edate) |> summarize(parties = n())
el_iterable <- seq_along(elections$parties)

for (e in el_iterable) {
  sub_df <- mp |> 
    filter(countryname == elections$countryname[e], 
           edate == elections$edate[e])
  center <- sub_df |> 
    summarize(across(per101:per706, mean))
  wgv_squared <- 0
  for (p in 1:elections$parties[e]) {
    party <- sub_df[p, 9:64]
    distance <- (party - center)^2 |> sum()
    wgv_squared <- wgv_squared + distance
  }
  elections$wgv[e] <- (wgv_squared / elections$parties[e])
}
```
Mutate elections

```{r}
elections <- elections |> rowwise() |>
  mutate(year = str_split(edate, "/")[[1]][3])
```


Export data to csv

```{r eval = FALSE}
write.csv(elections, "data/elections_export.csv", row.names = TRUE)
```


Exploratory analysis of WGV (partisan heterogeneity):
There seems to be a positive relationship between number of party in an election and my measure of party differences. Is this the result of poor operationalization or is it a substantive finding?

```{r}
#VIZ
elections |> ggplot(aes(x = parties, y = wgv)) + 
  geom_point() + geom_smooth(method = "lm")

elections |> group_by(parties) |> summarize(mean = mean(wgv), n = n()) |>
  ggplot(aes(x = parties, y = mean)) + geom_col() + 
  geom_text(aes(label = n), vjust = -0.2)
```

Effective number of parties
No of partie further than x from mean.
Weight dist by votes
Get map visualization

```{r}
#MAP VIZ
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

european_union <- c("Austria","Belgium","Bulgaria","Croatia","Cyprus",
                    "Czech Rep.","Denmark","Estonia","Finland","France",
                    "Germany","Greece","Hungary","Ireland","Italy","Latvia",
                    "Lithuania","Luxembourg","Malta","Netherlands","Poland",
                    "Portugal","Romania","Slovakia","Slovenia","Spain",
                    "Sweden","United Kingdom")

map <- ne_countries(scale = 50, returnclass = 'sf') |> 
  filter(name %in% european_union)
bbox_europe <- st_bbox(c(xmin = -10, ymin = 20, xmax = 50, ymax = 80), 
                       crs = st_crs(map))
elections_mean <- elections |> group_by(countryname) |>
  summarize(wgv = mean(wgv))
map_cropped <- st_crop(map, bbox_europe) |> 
  left_join(elections_mean, by = c("name" = "countryname"))

ggplot(data = map_cropped) +
  geom_sf(mapping = aes(fill = wgv)) +
  scale_fill_gradient(name = "Party differences", 
                      low = "#FF0000FF", high = "#FFFF00FF", 
                      na.value = "grey50") +
  labs(title = "Party heterogeneity in Europe") +
  theme(plot.title.position = "plot")

```

Loading & tidying electoral system data

```{r}
syst <- read_csv("data/elect_systems_national_leg_europe.csv") |> 
  filter(Country %in% countries)
names(syst) <- c("country", "isocode", "iso2166", "year", "elsys")
syst <- syst |> mutate(systfam = case_when(
  elsys == "List PR" ~ "Proportional Representation",
  elsys == "FPTP" | elsys == "STV" | elsys == "TRS" ~ "Plurality",
  elsys == "Parallel" | elsys == "MMP" ~ "Mixed"
  )) |>
  rowwise() |>
  mutate(year = toString(year))
```

Joining syst and elections

```{r}
#elections <- read_csv("data/elections_export.csv")
elections_sys <- elections |> 
  inner_join(syst, by = c("countryname" = "country", "year" = "year")) |>
  select(countryname, edate, year, parties, wgv, elsys, systfam)
elections_sys
```
Save election_sys

```{r}
write.csv(elections_sys, "data/elections_sys_export.csv", row.names = TRUE)
```


Linear regression

```{r}
fit <- lm(wgv ~ systfam, data = elections_sys)
summary(fit)
```

Visualizing differences

```{r}
syst_table <- elections_sys |> group_by(systfam) |>
  summarize(mean_wgv = mean(wgv)) 

knitr::kable(syst_table, digits = 1, col.names = c("System Family", "Average WGV"))

syst_table |> ggplot(aes(x = systfam, y = mean_wgv)) + 
  geom_col() + labs(title = "Average partisan distance by electoral system family",
                    x = "System family",
                    y = "Average WGV")
```

Statistical significance. Plurality vs Proportional Representation

```{r}
set.seed(02138)
diff <- elections_sys |> 
  filter(systfam %in% c("Proportional Representation", "Plurality")) |>
  specify(response = wgv, explanatory = systfam) |>
  calculate(stat = "diff in means", 
            order = c("Proportional Representation", "Plurality"))
diff

diff_null_dist <- elections_sys |>
  filter(systfam %in% c("Proportional Representation", "Plurality")) |>
  specify(response = wgv, explanatory = systfam) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", 
            order = c("Proportional Representation", "Plurality"))

plot_prop <- visualize(diff_null_dist) + 
  shade_p_value(obs_stat = diff, direction = "two-sided") + 
  labs(title = "Proportional Representation vs Plurality",
       x = "",
       y = "")

diff_p <- get_p_value(diff_null_dist, obs_stat = diff, direction = "two-sided")
diff_p
```
Statistical significance. Plurality vs Mixed

```{r}
set.seed(02138)
diff_mixed <- elections_sys |> 
  filter(systfam %in% c("Mixed", "Plurality")) |>
  specify(response = wgv, explanatory = systfam) |>
  calculate(stat = "diff in means", 
            order = c("Mixed", "Plurality"))
diff_mixed

diff_null_dist_mixed <- elections_sys |>
  filter(systfam %in% c("Mixed", "Plurality")) |>
  specify(response = wgv, explanatory = systfam) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", 
            order = c("Mixed", "Plurality"))

plot_mixed <- visualize(diff_null_dist_mixed) + 
  shade_p_value(obs_stat = diff_mixed, direction = "two-sided") + 
  labs(title = "Mixed vs Plurality",
       x = "",
       y = "")

diff_p_mixed <- get_p_value(diff_null_dist_mixed, obs_stat = diff_mixed, direction = "two-sided")
diff_p_mixed
```
Grid of plots

```{r include=FALSE}
png(file = "img/syst_diff_grid.png", height = 800, width = 1500)
gridExtra::grid.arrange(plot_prop, plot_mixed, ncol = 2)
dev.off()
```

